Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-08-06 surfaces-write-up/surface-write-up.tex
address bq^n -> bz^n todo on vector exp notation

In part, this work is motivated by a recent paper by
O'Dorney \cite{dorney:canonical}, which gives similar descriptions
of canonical rings for general $\bq$-divisors on $\bp^1$. This
study also has connections to the papers by Voight--Zureick-Brown 
\cite{vzb:stacky} and Landesman--Ruhm--Zhang \cite{lrz:spin-cring}
on canonical rings of stacky curves which provide bounds
on rings of modular forms of Fuchsian groups through the study
of particular $\bq$-divisors on curves.  \todo{Peter: I think this is too much specificity about our result.  I think everything after ``stacky curves'' should be eliminated, as it seems a little self-promoting.}

@@

In this subsection, we restrict attention to the case of effective
fractional divisors $D \in \bq \otimes \di \bp^m$. We completely
characterize the canonical rings when $D$ consists of one
hypersurface and give tight bounds on generator and relation
degrees for general effective divisors $D$.
\todo{Aaron: make a remark as to why these bounds are tight}
\begin{convention}
Let $[x_0, \ldots, x_m]$ be coordinates for $\bp^m$. Let
$\vec{v} = (v_0, \ldots, v_m) \in \bq^{m + 1}$. \todo{Peter: do we want to allow $\vec{v}\in \mathbb{Q}^{m+1}$ rather than $\mathbb{Z}^{m+1}$?  I'm not sure if the arguments actually need the generality (and if so, it is a little worrisome since fractional powers aren't clearly defined algebraically.

Aaron: No, these vectors should just lie in $\bz$. They're just used
to given an ordering on monomials.
}  Then write
\[
	x^{\vec{v}} := \prod_{i = 0}^{m} x_i^{v_i}
\]
\end{convention}

\begin{defn}
\label{defn:vec-sum}
For $\vec{v} \in \bq^n$, denote $\deg \vec{v} := \sum_{i = 0}
^n v_i$.
Let 
\begin{align*}
	\mss_i := \left \{\vec{v} \in \bz_{\geq 0}^{m + 1} \; : \;
\deg \vec v = c_i \right\}.	
\end{align*}

@@

\begin{defn}
\label{defn:vec-order}
Let $\vec{v}, \vec{w} \in \bq^n$. Let $i \in \{1,\ldots, n\}$
be the biggest index such that $v_i$ is nonzero
and $j \in \{1,\ldots, n\}$ be the smallest index such that $w_j$ is
nonzero. Define $\vec{v} \prec \vec{w}$ if $i \leq j$.
\end{defn}

>>>>

In part, this work is motivated by a recent paper by
O'Dorney \cite{dorney:canonical}, which gives similar descriptions
of canonical rings for general $\bq$-divisors on $\bp^1$. This
study also has connections to the papers by Voight--Zureick-Brown 
\cite{vzb:stacky} and Landesman--Ruhm--Zhang \cite{lrz:spin-cring}
on canonical rings of stacky curves which provide bounds
on rings of modular forms of Fuchsian groups through the study
of particular $\bq$-divisors on curves.  \todo{I think this is too much 
specificity about our result.  I think everything after ``stacky curves'' should be 
eliminated, as it seems a little self-promoting.}

@@

In this subsection, we restrict attention to the case of effective
fractional divisors $D \in \bq \otimes \di \bp^m$. We completely
characterize the canonical rings when $D$ consists of one
hypersurface and give tight bounds on generator and relation
degrees for general effective divisors $D$.
\todo{Aaron: make a remark as to why these bounds are tight}
\begin{convention}
Let $[x_0, \ldots, x_m]$ be coordinates for $\bp^m$. Let
$\vec{v} = (v_0, \ldots, v_m) \in \bz^{m + 1}$.  Then write
\[
	x^{\vec{v}} := \prod_{i = 0}^{m} x_i^{v_i}
\]
\end{convention}

\begin{defn}
\label{defn:vec-sum}
For $\vec{v} \in \bz^n$, denote $\deg \vec{v} := \sum_{i = 0}
^n v_i$.
Let 
\begin{align*}
	\mss_i := \left \{\vec{v} \in \bz_{\geq 0}^{m + 1} \; : \;
\deg \vec v = c_i \right\}.	
\end{align*}

@@

\begin{defn}
\label{defn:vec-order}
Let $\vec{v}, \vec{w} \in \bz^n$. Let $i \in \{1,\ldots, n\}$
be the biggest index such that $v_i$ is nonzero
and $j \in \{1,\ldots, n\}$ be the smallest index such that $w_j$ is
nonzero. Define $\vec{v} \prec \vec{w}$ if $i \leq j$.
\end{defn}
